---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# cordis

<!-- badges: start -->
<!-- badges: end -->

The goal of `cordis` is to simplify data access to data from [CORDIS](https://cordis.europa.eu/), which is an acronym for the Community Research and Development Information Service. It is the European Commission's primary source of results from the projects funded by the EU's framework programmes for research and innovation. This includes programmes from FP1 to Horizon 2020. 

## Data preparation

CORDIS makes data about European research projects available at various locations, such as https://data.europa.eu/euodp/en/data/dataset/cordisH2020projects. The download speed is rate limited when working directly against these files, and file formats and compression are different. With some data preparation, these datasets can be loaded into a database, providing simplified and faster local access to the data.

In [`data-raw`](data-raw) there are data preparation scripts which download data from these locations into a local cache directory (`~/.cache/cordis`). A [duckdb](https://duckdb.org) database is built from these files and uploaded using [piggyback](https://github.com/ropensci/piggyback) to a GitHub repo at https://github.com/KTH-Library/cordis-data as a versioned GitHub Release. 

There is also a function for exporting the entire database in Parquet format, which allows moving the data for example to a [Minio](https://min.io/) server, where it can be accessed by other data integration tools like Apache Spark. Only package developers need to use these functions when preparing and updating the data.

## Usage

The data in the "github releases" location mentioned above can be installed locally by regular package users by running "cordis_import()", a function for importing the data from the data repository. This needs to be done once. The download and upload rate is good.

Users can then make a connection to the database locally. This allows arbitrary in-process data processing with tidyverse tools such as dplyr.

Convenience functions allows for inspecting the database schema / finding table and field names. 

## Installation

You can install the released version of `cordis` from [GitHub](https://github.com/KTH-Library/cordis) with:

``` r
devtools::install_github("KTH-Library/cordis")
# run once to install local data
cordis_import()
```

## Example

This is a basic example which shows you how to work with the data:

```{r example}
library(cordis)
suppressPackageStartupMessages(library(dplyr))
library(knitr)

# tables in the database
cordis_tables() %>%
  arrange(desc(n_row))

# database schema
cordis_schema() %>%
  head(20)

# get a connection
con <- cordis_con()
# remember to disconnect when done:
# cordis_disconnect(con)

# display first five rows with PI data
con %>% tbl("pi") %>% 
  head(5) %>% 
  knitr::kable()

# display first row of projects info
con %>% tbl("projects") %>% 
  head(1) %>% 
  glimpse()

# display first row with publications data
con %>% tbl("projectpublications") %>% 
  head(1) %>% 
  select(-starts_with("X")) %>% 
  glimpse()

cordis_disconnect(con)

```

